---
title: "Biological control of important fungal diseases of potato and raspberry by two *Bacillus velezensis* strains"
author: "Lourdes Arellano M, Antonia Fernandes, Juliana Cajiao and Rosario Heribia"
date: "22/12/2021"
output:
  html_document:
    code_folding: hide
---
## Introduction 

Fungal phytopathogens affecting agricultural crops lead to a decrease in their quality and production. They act as a threat to crops through various mechanisms of pathogenesis that compromise the immune system of the plants. The use of fungicides against fungal plant diseases improves crop yield, quality, and shelf-life (Brauer et al., 2019).  Nevertheless, the irrational uses of this products triggers to health problems, and can also cause environmental problems. Accumulation of pesticides in the environment disrupts the ecological balance and gives rise to pathogenic resistance to the pesticides. Thus, the application of fungicides in agriculture should be under strict regulation to ensure that commercialized foods are safe for consumption and pose negligible risks of acute toxicity due to carry-over (Brauer et al., 2019).   

In the case of potato and raspberry plants, they are affected by important diseases among which are mainly stem canker and black scurf, causes by the fungal Rhizoctonia solani and Didymella applanata, namely. Recently, two possible strains of the bacteria Bacillus velezensis, specifically BZR336 g and BZR 517, have been demonstrated as potential biological control of these diseases (Asaturova et al., 2020).

Bacillus velezensis, an antagonistic bacterium produces several antimicrobial compounds as secondary metabolites against various phytopathogens. B. velezensis produces spores, which can be used for the development of an effective microbial biopesticide formulation. Previously, many studies have demonstrated that several secondary metabolites produced by antagonistic bacteria, play key roles in the control of various phytopathogens. Furthermore, Bacillus spp., have been found to produce cyclic lipopeptides (CLPs), which play a key role in antimicrobial activity (Soo Kim et al., 2021).  

Different methodologies can be applied for the evaluation of the effectiveness of pesticides with different factors and levels; and for this, a variance test must first be carried out. An example is Duncan's test, which is used to compare all pairs of means. It was first developed by Duncan in 1951 but later he himself modified his first method generating what is now called Duncan's New Multiple Range method.

Therefore, in its equation, this test implicitly measures the number of inclusive means between the two means to compare for balanced designs. To apply this test, the following stages must be passed: determine the standard error (standard deviation) of each mean; obtained from the Anova table. Then, with the degrees of freedom of the error and the level of significance, the values, intervals or standardized significant amplitudes are determined using Duncan's tables of standardized amplitudes, then the minimum significant amplitudes are determined. Following this, the average results of the experiment are ordered in increasing order and the ordered means are compared: The highest average, with the lowest, comparing this difference with the least significant interval.

This process continues until the differences between all possible pairs have been compared. If an observed difference is greater than the minimum significant interval, it is concluded that the pair of means compared are significantly different (UNAL, 2021). 

Another variance test that can be applied is the Tukey test. This method was proposed by Tukey (1952) and is very similar in application to the DMS test, except for the fact that instead of using the  distributions as the basis for making comparisons, the standardized or studentized range distribution is used. In this test, a single value is used against which all possible pairs of means are compared. Tukey's comparison method was reformed by Kramer (1956) for cases in which the number of replicas is not equal. This method is known as the Tukey-Kramer method. (UNAL, 2021).

By the other hand, the Shapiro-Wilks test poses the null hypothesis that a sample comes from a normal distribution.  The Shapiro-Wilks test poses the null hypothesis that a sample comes from a normal distribution. This test attempts to reject the null hypothesis according to the level of significance. Therefore, if the significance is greater than 0.05, the null hypothesis is overcome, and if it is less than this value, the alternate hypothesis is accepted and it could then be stated that the data do not follow a normal distribution (UNAL, 2021).
It is important to highlight that the study of beneficial microorganisms to create new biological formulations for plant protection is typically conducted under the same ecological and geographical conditions in which the microorganisms were isolated. Thus, effect of this bacterium has been demonstrated as a biological control in crops such as tomato and in specific conditions. However, its potential use as a pesticide in potatoes and raspberries is unknown. That said, the main objective of this work was to demonstrate the potential effect of B. velezensis strains against R. solani, as a biological control in raspeberry and potato, and its potential effect on potato growth increseament, and the strains were isolated in southern Russia on suppressing Rhizoctonia disease on potato and spur blight on red raspberry in Western Siberia, which is a geographically distant region with much more severe climate conditions (Asaturova et al., 2020).

For this, different experiments and scenarios were carried out. Thus, the potential effect was measured through the antifungal activity of B. velezensis against R. solani in vitro, by applying different treatments in different proportions (quantity) of colony units. The evaluation of the effect of B. velezensis strains on the incidence and severity of the disease was evaluated through morphometric parameters caused by R. solani in two potato cultivars differentiated by resistance to the disease, during the growing seasons. 2015 and 2016 in Western Siberia. Finally, to test the potential increase in potato growth, the biomass distribution of the daughter tubers of two potato cultivars and the total yield were measured during the 2015 and 2016 growing seasons in Western Siberia.

The diameter of the colonies (cm) was measured in R. solani and D. applanata under the application of the strains: Bacillus velezensis BZR 336 g and BZR 517 at three different concentrations: 104, 105 and 106 CFU / ml and on three different days : 3, 5 and 7. Disease incidence and severity was measured in percentage in the years 2015 and 2016, in three different treatments: control, BZR 336 g strain and BZR 517 strain, at weeks 4, 6 and 10 after sowing.
Morphometric parameters of two different cultivars (Svitanok kievskiy and Yuna) were measured, such as the total height of the plants (cm) and the number of stems developed with the same factors and levels previously mentioned.

In these same two cultivars the biomass distribution by fractions was measured, in order to measure the plant grow capacity, and they were divided in tree fraction categories: small (less than 35 g), medium (36–180 g), or large ( more than 180 g). Yields were determined by weighing all daughter tubers from each plot in the both years. 
For statistical tests, normality and variance distribution tests were applied, such as Shapiro-Wilks, and for significance between Duncan and Tukey tests.


## Objectives

Demonstrate the potential effect of *B. velezensis* strains  against *R. solani* as a biological control, and a potential increasement in the plant growth of potato. 

Also in this review, the aim of this group is to compare the analysis done in the article between the Duncan and Turkey tests to understand why Duncan was used in the article.


## Hypothesis

*B. velezensis* strains can be used as biological control agents against *R. solani*, and increase significantly the plant growth of potato. Moreover, there is no significant difference between the results obtained through the Duncan and Turkey test.

## Methods


**1. Dowload Data:**
First, dowload the raw data of the paper that is in the link: 
<https://dfzljdn9uc3pi.cloudfront.net/2021/11578/1/peerj-55829-Raw_data.xls>

**Article:**  <https://peerj.com/articles/11578/#>

```{r}
if(!dir.exists("data")) dir.create("data")
download.file("https://dfzljdn9uc3pi.cloudfront.net/2021/11578/1/peerj-55829-Raw_data.xls",
              "data/rawdata.xls")
```

**2. Open and Arrange Data:** For this step, it was used the package "Readxl" to charge the excel file. Each sheet of excel was opened indicating its name and the range required. In some sheets it was use the function "fill" to fill the cells that were empty. Also some names of variables were changed with "rename" and "pivot longer" was used to increase the number of rows or to decrease the number of columns.The potence format was changed. Finally, some new sheets were created why others where joined or deleted in order to have the data arranged. 

```{r include = FALSE}
library(readxl)
readxl::excel_sheets("data/peerj-55829-Raw_data.xls")
dir("data")
#dir("C:\Users\lsare\Desktop\Plant Health\Estadistica\Biologicalcontrol\data\")
library(readxl)
#AntagonisticActivity <- read_excel("data/peerj-55829-Raw_data.xls")
```


```{r include= FALSE}
#Open sheets with ranges 
library
AntagonisticActivity <-
  read_excel("data/peerj-55829-Raw_data.xls",
             sheet = "Antagonistic activity",
             range = "A3:E38")
DiseaseIncidenceAndSeverity <- read_excel("data/peerj-55829-Raw_data.xls",
                                          sheet = "disease incidence and severity",
                                          range = "A4:N40")
MorphometricCharactOfPotato <- read_excel("data/peerj-55829-Raw_data.xls",
                                          sheet = "morphometric charact of potato",)
BiomassDistribution <- read_excel("data/peerj-55829-Raw_data.xls",sheet = "biomass distribution")
YieldOfPotato <- read_excel("data/peerj-55829-Raw_data.xls",
                            sheet = "yield of potato",
                            range = "B3:I15")
#Fill 
library(tidyr)
library(dplyr)
AntagonisticActivity <- AntagonisticActivity %>% fill(everything(), .direction = "down")
BiomassDistribution <- BiomassDistribution %>% fill(everything(), .direction = "down")
DiseaseIncidenceAndSeverity <- DiseaseIncidenceAndSeverity %>% fill(everything(), .direction = "down")
MorphometricCharactOfPotato <- MorphometricCharactOfPotato %>% fill(everything(), .direction = "down")
YieldOfPotato <- YieldOfPotato %>% fill(everything(), .direction = "down")

#Rename 
library(dplyr)
names(AntagonisticActivity)[names(AntagonisticActivity) == 'strain'] <- "Strain"
names(AntagonisticActivity)[names(AntagonisticActivity) == 'CFU/ml'] <- "Concentration"

#Pivot longer 
AntagonisticActivity <- pivot_longer(AntagonisticActivity,cols= c("3d", "5d", "7d"), names_to="Days", values_to="Colony.Diameter")

#change potence
AntagonisticActivity$Concentration
a <- as.numeric(substr(AntagonisticActivity$Concentration, 1, 2))
b <- as.numeric(substr(AntagonisticActivity$Concentration, 3, 3))
a^b
AntagonisticActivity$Concentration <- a^b
rm(a,b)

#Create new column for treatments
AntagonisticActivity$Trat <- paste(AntagonisticActivity$Strain,AntagonisticActivity$Concentration)

#Rename days
library(dplyr)
AntagonisticActivity$Days<-gsub("d","",as.character(AntagonisticActivity$Days))

#Join data frame 
library(dplyr)
Data.Cultivars <- full_join(MorphometricCharactOfPotato, DiseaseIncidenceAndSeverity, by = c("Cultivar","Year", "Treatment","Weeks after planting"))

#Rename 
library(dplyr)
names(Data.Cultivars)[names(Data.Cultivars) == 'Plant height, cm'] <- "1"
names(Data.Cultivars)[names(Data.Cultivars) == '...6.x'] <- "2"
names(Data.Cultivars)[names(Data.Cultivars) == '...7.x'] <- "3"
names(Data.Cultivars)[names(Data.Cultivars) == '...8.x'] <- "4"
names(Data.Cultivars)[names(Data.Cultivars) == '...9.x'] <- "5"

#Pivot longer 
Data.Cultivars <- pivot_longer(Data.Cultivars,cols= c("1", "2", "3","4","5"), names_to="Rep", values_to="Plant.Height")

#Create empty new columns 
Data.Cultivars$Disease.Incidence <- NA
Data.Cultivars$Disease.Severity <- NA
Data.Cultivars$Stems <- NA

#Add data to the columns 
variables <- c("Disease incidence, %","...6.y","...7.y","...8.y","...9.y")
for (i in 1:5) {
  A <- Data.Cultivars$Rep==i
  Data.Cultivars$Disease.Incidence[A] <- Data.Cultivars[[variables[i]]][A]
}

Data.Cultivars$Stems <- NA
variables1 <- c("The number of stems per plant","...11.x","...12.x","...13.x","...14.x")
for (i in 1:5) {
  A1 <- Data.Cultivars$Rep==i
  Data.Cultivars$Stems[A1] <- Data.Cultivars[[variables1[i]]][A1]
}

Data.Cultivars$Disease.Severity <- NA
variables1 <- c("Disease severity, %","...11.y","...12.y","...13.y","...14.y")
for (i in 1:5) {
  A2 <- Data.Cultivars$Rep==i
  Data.Cultivars$Disease.Severity[A2] <- Data.Cultivars[[variables1[i]]][A2]
}


#dataframe with variables that I need 
namesdatacultivar <- c("Cultivar","Treatment","Year","Weeks after planting", "Rep","Plant.Height","Disease.Incidence","Disease.Severity","Stems")           
Data.Cultivars.Final <- Data.Cultivars [which(names(Data.Cultivars) %in% namesdatacultivar)]

#Yieldofpotato 
#Rename 
library(dplyr)
names(YieldOfPotato)[names(YieldOfPotato) == 'Yield, t ha-1'] <- "1"
names(YieldOfPotato)[names(YieldOfPotato) == '...5'] <- "2"
names(YieldOfPotato)[names(YieldOfPotato) == '...6'] <- "3"
names(YieldOfPotato)[names(YieldOfPotato) == '...7'] <- "4"
names(YieldOfPotato)[names(YieldOfPotato) == '...8'] <- "5"
#Pivot longer 
YieldOfPotato <- pivot_longer(YieldOfPotato,cols= c("1", "2", "3","4","5"), names_to="Rep", values_to="Yield.Potato")

#BiomassDistribution
#Rename 
library(dplyr)
names(BiomassDistribution)[names(BiomassDistribution) == 'Biomass distribution by fractions, %'] <- "1"
names(BiomassDistribution)[names(BiomassDistribution) == '...6'] <- "2"
names(BiomassDistribution)[names(BiomassDistribution) == '...7'] <- "3"
names(BiomassDistribution)[names(BiomassDistribution) == '...8'] <- "4"
names(BiomassDistribution)[names(BiomassDistribution) == '...9'] <- "5"
#Pivot longer 
BiomassDistribution <- pivot_longer(BiomassDistribution,cols= c("1", "2", "3","4","5"), names_to="Rep", values_to="Percentage.Biomass")

#create a sheet for medium fraction
medium.fraction <- BiomassDistribution[BiomassDistribution$Fractions=="medium",]

#join sheets of yield and medium fraction
names(YieldOfPotato)[names(YieldOfPotato) == 'Bacillus strain'] <- "Treatment"
YieldOfPotato <- full_join(YieldOfPotato,medium.fraction)
YieldOfPotato <- YieldOfPotato[,-which(names(YieldOfPotato)=="Fractions")]

#remove sheets
ls()
rm("Data.Cultivars","DiseaseIncidenceAndSeverity","medium.fraction","MorphometricCharactOfPotato")

```


**3. Do analysis:**
Finally with the arrangement of the sheets of excel, there were obtained 4 main data sets. The first one was Antagonistic Activity, then Data Cultivars that include the variables of Percentage of Disease Incidence, Percentage of Disease Severity, Plant height and number of stems per plant.Then, the data set of Biomass Distribution and finally the data set of Yield of Potato. 

With these data sets were performed Tukey HSD Test, Duncan test and Normality Test of Residuals with shapiro Test. All results were expressed in graphs using the packages of "ggplot" and "ggpubr". The results were analyzed and compared with the results of the article. It is important to say that in the article, the authors have only performed Duncan Test.

**Description of data sets:**

**1) Effect of *B. velezensis* strains on disease incidence and severity and morphometric parameters caused by *R. solani* on two potato cultivars differed by disease resistance during 2015 and 2016 growing seasons in Western Siberia**


Cultivars:

1. Svitanok kievskiy 

2. Yuna


Years:

* 2015

* 2016


Treatments:

1. Control

2. BZR 336g 

3. BZR 517


Weeks after planting:

* 4

* 6

* 10


Variables: (5 rep/treatment)

1. Disease incidence (%) 

2. Disease severity (%)

3. Plant height (cm) 

4. Number of stems per plant


&nbsp;

**2) Effect of *B. velezens* strains on the biomass distribution of daughter tubers of two potato cultivars differed by disease resistance to *R. solani* and total yield during 2015 and 2016 growing seasons in Western Siberia.**


Cultivars:

1. Svitanok kievskiy 

2. Yuna


Years: 

* 2015

* 2016


Treatments:

1. Control 

2. BZR 336g

3. BZR 517


Fractions:

* Small

* Medium

* Large 


Variables: (5 rep/treatment)

1. Biomass distribution 

**3) Antifungal activity of B. velezensis against *R solani* in vitro.** 

Strains:

1. *Bacillus velezensis* strain BZR 336 g

2. *Bacillus velezensis* strain BZR 517

3. Control 


Treatments:
 
1. 10^4 CFU/mL
  
2. 10^5 CFU/mL
  
3. 10^6 CFU/mL
  
4. Control.

###### CFU: Colony-Forming Unit


Days:

* 3

* 5

* 7 

Variables: 

* Colony Diameter (cm)


&nbsp;

## Results and Discussion 


&nbsp;

**1) Effect of *B. velezensis* strains on disease incidence and severity and morphometric parameters caused by *R. solani* on two potato cultivars differed by disease resistance during 2015 and 2016 growing seasons in Western Siberia**


&nbsp;
**1.1 Plant Height**

The Duncan and Tukey tests had few differences.Only in year 2016, week 6 in cultivar Svitanok kievski it can be seen a difference in the group of BZR 517 (Figures 1 and 2). The Shapiro-Wilk normality test was performed for Anova and Square Root Anova because in Anova p was lower than 0.05, that means that the data significantly deviate from a normal distribution (Figure 3).

Taking into account the plant height, at week 10, we can see that the BZR 336g treatment generated greater plant height in every cultivar and year in the Ducan test, less in the cultivar Svitanok Kievski in 2016, when there was no significant difference between treatments. In the Tukey test, the difference between treatments was significant only in the Yuna variety.


```{r include = FALSE}
#Duncan Analysis Data.Cultivars.Final
library(agricolae)
library(ggplot2)
library(ggpubr)

# Plant Height

#Anova Plant Height
model<-aov(Plant.Height~Treatment+`Weeks after planting`+Year, data=Data.Cultivars.Final)
summary(model)
Data.Cultivars.Final$`Weeks after planting` <- factor(Data.Cultivars.Final$`Weeks after planting`)

## Plant Height Duncan 

Y <- W <- Tr <- L <- C <- M <- numeric()

for (i in 2015:2016) { 
  for (k in levels(as.factor(Data.Cultivars.Final$Cultivar))) {
    for(j in levels(Data.Cultivars.Final$`Weeks after planting`)) {
      d1 <- Data.Cultivars.Final[Data.Cultivars.Final$Year==i & 
                                   Data.Cultivars.Final$`Weeks after planting`==j &
                                   Data.Cultivars.Final$Cultivar==k, ]
      model <- aov(Plant.Height~Treatment, data=d1)
      summary(model)
      DT <- duncan.test(model, "Treatment", console = TRUE)
      Y <- c(Y, i,i,i)
      W <- c(W, j,j,j)
      C <- c(C, k,k,k)
      Tr <- c(Tr, row.names(DT$groups))
      L <- c(L, DT$groups$groups)
      M <- c(M, DT$groups$Plant.Height)
    } 
  }
}


LettersPH <- data.frame(Year = Y, `Weeks after planting` = factor(W), 
                        Cultivar = C, Treatment = factor(Tr), letra = L, mea = M)

## Plant Height Tukey 

Y <- W <- Tr <- L <- C <- M <- numeric()

for (i in 2015:2016) { 
  for (k in levels(as.factor(Data.Cultivars.Final$Cultivar))) {
    for(j in levels(Data.Cultivars.Final$`Weeks after planting`)) {
      d1z <- Data.Cultivars.Final[Data.Cultivars.Final$Year==i & 
                                    Data.Cultivars.Final$`Weeks after planting`==j &
                                    Data.Cultivars.Final$Cultivar==k, ]
      model <- aov(Plant.Height~Treatment, data=d1z)
      summary(model)
      DT <- HSD.test(model, "Treatment", console = TRUE)
      Y <- c(Y, i,i,i)
      W <- c(W, j,j,j)
      C <- c(C, k,k,k)
      Tr <- c(Tr, row.names(DT$groups))
      L <- c(L, DT$groups$groups)
      M <- c(M, DT$groups$Plant.Height)
    } 
  }
}

LettersPHz <- data.frame(Year = Y, `Weeks after planting` = factor(W), 
                         Cultivar = C, Treatment = factor(Tr), letra = L, mea = M)

```

```{r}
library(agricolae)
library(ggplot2)
library(ggpubr)
#Plant Height Duncan Graph

ggplot(Data.Cultivars.Final, aes(x=`Weeks after planting`, y=Plant.Height, fill=Treatment))+ylab("Plant Height (cm)")+
  geom_boxplot()+
  facet_grid(Cultivar~Year)+
  theme_light()+
  geom_text(data = LettersPH,
            mapping = aes(x = Weeks.after.planting,
                          y = mea + 5,
                          label = letra),
            position = position_dodge(0.9))+
  labs(caption = "Figure No 1. Results of Duncan Test in the variable of Plant Height on potato (cm)") +
  theme(plot.caption.position = "plot",
        plot.caption = element_text(hjust = 0))
```

```{r}
#Plant Height Tukey Graph

ggplot(Data.Cultivars.Final, aes(x=`Weeks after planting`, y=Plant.Height, fill=Treatment))+ylab("Plant Height (cm)")+
  geom_boxplot()+
  facet_grid(Cultivar~Year)+
  theme_light()+
  geom_text(data = LettersPHz,
            mapping = aes(x = Weeks.after.planting,
                          y = mea + 5,
                          label = letra),
            position = position_dodge(0.9))+
  labs(caption = "Figure No 2. Results of Tukey Test in the variable of Plant Height on potato (cm)") +
  theme(plot.caption.position = "plot",
        plot.caption = element_text(hjust = 0))

#Residuals Plant Height 
anova.ph<-aov(Plant.Height~Treatment+`Weeks after planting`+Year, data=Data.Cultivars.Final) #anova 
anova.ph.sqrt <- aov(sqrt(Plant.Height)~Treatment+`Weeks after planting`+Year, data=Data.Cultivars.Final) #square root 
shapiro.test(anova.ph$residuals) #0.0329 anormal 
shapiro.test(anova.ph.sqrt$residuals) #0.02933
par(mfrow = c(1, 2), oma = c(0, 0, 1.1, 0))
plot(anova.ph, main = "ANOVA", which = 2, ask = F,sub.caption = "")
plot(anova.ph.sqrt, main = "Square Root ANOVA", which = 2, ask = F,sub.caption = "")
```

<font size="1.5"> Figure No 3. Normality of Residuals Plant Height.</font> 



&nbsp;

**1.2 Disease Incidence** 

The Duncan and Tukey tests had few differences.Only in year 2015 in cultivar Svitanok kievski it can be seen a difference between the groups (Figures 4 and 5). The Shapiro-Wilk normality test was performed for Anova and Square Root Anova because in Anova p was lower than 0.05, that means that the data significantly deviate from a normal distribution (Figure 6). 

Taking into account the incidence of the disease, at week 10, we can observe that the BZR 336 treatment was the one that caused the lowest incidence in both years and in both cultivars when using the Duncan test. In the Tukey test, this difference was not significant between treatments on cultivar Svitanok kievski in 2015.


```{r include = FALSE}
#Disease Incidence
#Anova Disease Incidence 
model<-aov(Disease.Incidence~Treatment+`Weeks after planting`+Year, data=Data.Cultivars.Final)
summary(model)
Data.Cultivars.Final$`Weeks after planting` <- factor(Data.Cultivars.Final$`Weeks after planting`)


#Duncan Test Disease Incidence

Y <- W <- Tr <- L <- C <- M <- numeric()

for (i in 2015:2016) { 
  for (k in levels(as.factor(Data.Cultivars.Final$Cultivar))) {
    for(j in levels(Data.Cultivars.Final$`Weeks after planting`)) {
      d1 <- Data.Cultivars.Final[Data.Cultivars.Final$Year==i & 
                                   Data.Cultivars.Final$`Weeks after planting`==j &
                                   Data.Cultivars.Final$Cultivar==k, ]
      model <- aov(Disease.Incidence~Treatment, data=d1)
      summary(model)
      DT <- duncan.test(model, "Treatment", console = TRUE)
      Y <- c(Y, i,i,i)
      W <- c(W, j,j,j)
      C <- c(C, k,k,k)
      Tr <- c(Tr, row.names(DT$groups))
      L <- c(L, DT$groups$groups)
      M <- c(M, DT$groups$Disease.Incidence)
    } 
  }
}


LettersDI <- data.frame(Year = Y, `Weeks after planting` = factor(W), 
                        Cultivar = C, Treatment = factor(Tr), letra = L, mea = M)

#Tukey Disease Incidence

Y <- W <- Tr <- L <- C <- M <- numeric()

for (i in 2015:2016) { 
  for (k in levels(as.factor(Data.Cultivars.Final$Cultivar))) {
    for(j in levels(Data.Cultivars.Final$`Weeks after planting`)) {
      d1z <- Data.Cultivars.Final[Data.Cultivars.Final$Year==i & 
                                    Data.Cultivars.Final$`Weeks after planting`==j &
                                    Data.Cultivars.Final$Cultivar==k, ]
      model <- aov(Disease.Incidence~Treatment, data=d1z)
      summary(model)
      DT <- HSD.test(model, "Treatment", console = TRUE)
      Y <- c(Y, i,i,i)
      W <- c(W, j,j,j)
      C <- c(C, k,k,k)
      Tr <- c(Tr, row.names(DT$groups))
      L <- c(L, DT$groups$groups)
      M <- c(M, DT$groups$Disease.Incidence)
    } 
  }
}


LettersDIz <- data.frame(Year = Y, `Weeks after planting` = factor(W), 
                         Cultivar = C, Treatment = factor(Tr), letra = L, mea = M)
```

```{r}
#Duncan Graph Disease Incidence
ggplot(Data.Cultivars.Final, aes(x=`Weeks after planting`, y=Disease.Incidence, fill=Treatment))+ylab("Disease Incidence(%)")+
  geom_boxplot()+
  facet_grid(Cultivar~Year)+
  theme_light()+
  geom_text(data = LettersDI,
            mapping = aes(x = Weeks.after.planting,
                          y = mea + 8,
                          label = letra),
            position = position_dodge(0.9))+
  labs(caption = "Figure No 4. Results of Duncan Test in the variable of Disease Incidence (%)") +
  theme(plot.caption.position = "plot",
        plot.caption = element_text(hjust = 0))

#Tukey Graph Disease Incidence
ggplot(Data.Cultivars.Final, aes(x=`Weeks after planting`, y=Disease.Incidence, fill=Treatment))+ylab("Disease Incidence(%)")+
  geom_boxplot()+
  facet_grid(Cultivar~Year)+
  theme_light()+
  geom_text(data = LettersDIz,
            mapping = aes(x = Weeks.after.planting,
                          y = mea + 8,
                          label = letra),
            position = position_dodge(0.9))+
  labs(caption = "Figure No 5. Results of Tukey Test in the variable of Disease Incidence (%)") +
  theme(plot.caption.position = "plot",
        plot.caption = element_text(hjust = 0))
#Residuals Disease Incidence
anova.di<-aov(Disease.Incidence~Treatment+`Weeks after planting`+Year, data=Data.Cultivars.Final) #anova 
anova.di.sqrt <- aov(sqrt(Disease.Incidence)~Treatment+`Weeks after planting`+Year, data=Data.Cultivars.Final) #square root 
shapiro.test(anova.di$residuals) #0.03132 anormal 
shapiro.test(anova.di.sqrt$residuals) #0.001165
par(mfrow = c(1, 2), oma = c(0, 0, 1.1, 0))
plot(anova.di, main = "ANOVA", which = 2, ask = F,sub.caption = "")
plot(anova.di.sqrt, main = "Square Root ANOVA", which = 2, ask = F,sub.caption = "")
```

<font size="1"> Figure No. 6 Normality of Residuals Disease Incidence</font> 


**1.3 Disease Severity**

The Duncan and Tukey tests did not have differences (Figures 7 and 8). The Shapiro-Wilk normality test was performed for Anova and Square Root Anova because in Anova p was lower than 0.05, that means that the data significantly deviate from a normal distribution (Figure 9).

Taking into account the disease severity, at week 10, we can observe that the BZR 336 treatment was the one that caused the lowest severity in both years and in both cultivars in both tests. However, this difference was not significant between the two treatments on cultivar Svitanok kievski in 2015, in both tests.

```{r include = FALSE}
#Disease Severity

#Anova Disease Severity
model<-aov(Disease.Severity~Treatment+`Weeks after planting`+Year, data=Data.Cultivars.Final)
summary(model)
Data.Cultivars.Final$`Weeks after planting` <- factor(Data.Cultivars.Final$`Weeks after planting`)
#Disease Severity Duncan Test

Y <- W <- Tr <- L <- C <- M <- numeric()

for (i in 2015:2016) { 
  for (k in levels(as.factor(Data.Cultivars.Final$Cultivar))) {
    for(j in levels(Data.Cultivars.Final$`Weeks after planting`)) {
      d1 <- Data.Cultivars.Final[Data.Cultivars.Final$Year==i & 
                                   Data.Cultivars.Final$`Weeks after planting`==j &
                                   Data.Cultivars.Final$Cultivar==k, ]
      model <- aov(Disease.Severity~Treatment, data=d1)
      summary(model)
      DT <- duncan.test(model, "Treatment", console = TRUE)
      Y <- c(Y, i,i,i)
      W <- c(W, j,j,j)
      C <- c(C, k,k,k)
      Tr <- c(Tr, row.names(DT$groups))
      L <- c(L, DT$groups$groups)
      M <- c(M, DT$groups$Disease.Severity)
    } 
  }
}


LettersDS <- data.frame(Year = Y, `Weeks after planting` = factor(W), 
                        Cultivar = C, Treatment = factor(Tr), letra = L, mea = M)

#Tukey Test Disease Severity

Y <- W <- Tr <- L <- C <- M <- numeric()

for (i in 2015:2016) { 
  for (k in levels(as.factor(Data.Cultivars.Final$Cultivar))) {
    for(j in levels(Data.Cultivars.Final$`Weeks after planting`)) {
      d1z <- Data.Cultivars.Final[Data.Cultivars.Final$Year==i & 
                                    Data.Cultivars.Final$`Weeks after planting`==j &
                                    Data.Cultivars.Final$Cultivar==k, ]
      model <- aov(Disease.Severity~Treatment, data=d1z)
      summary(model)
      DT <- HSD.test(model, "Treatment", console = TRUE)
      Y <- c(Y, i,i,i)
      W <- c(W, j,j,j)
      C <- c(C, k,k,k)
      Tr <- c(Tr, row.names(DT$groups))
      L <- c(L, DT$groups$groups)
      M <- c(M, DT$groups$Disease.Severity)
    } 
  }
}


LettersDSz <- data.frame(Year = Y, `Weeks after planting` = factor(W), 
                         Cultivar = C, Treatment = factor(Tr), letra = L, mea = M)
```

```{r}
#Duncan Graph Disease Severity
ggplot(Data.Cultivars.Final, aes(x=`Weeks after planting`, y=Disease.Severity, fill=Treatment))+ylab("Disease Severity (%)")+
  geom_boxplot()+
  facet_grid(Cultivar~Year)+
  theme_light()+
  geom_text(data = LettersDS,
            mapping = aes(x = Weeks.after.planting,
                          y = mea + 5,
                          label = letra),
            position = position_dodge(0.9))+
  labs(caption = "Figure No 7. Results of Duncan Test in the variable of Disease Severity (%)") +
  theme(plot.caption.position = "plot",
        plot.caption = element_text(hjust = 0))

#Tukey Graph Disease Severity

ggplot(Data.Cultivars.Final, aes(x=`Weeks after planting`, y=Disease.Severity, fill=Treatment))+ylab("Disease Severity(%)")+
  geom_boxplot()+
  facet_grid(Cultivar~Year)+
  theme_light()+ 
  geom_text(data = LettersDSz,
            mapping = aes(x = Weeks.after.planting,
                          y = mea + 7,
                          label = letra),
            position = position_dodge(0.9))+
  labs(caption = "Figure No 8. Results of Tukey Test in the variable of Disease Severity (%)") +
  theme(plot.caption.position = "plot",
        plot.caption = element_text(hjust = 0))
#Residuals Disease Severity
anova.ds<-aov(Disease.Severity~Treatment+`Weeks after planting`+Year, data=Data.Cultivars.Final) #anova 
anova.ds.sqrt <- aov(sqrt(Disease.Severity)~Treatment+`Weeks after planting`+Year, data=Data.Cultivars.Final) #square root 
shapiro.test(anova.ds$residuals) #0.002356 anormal  
shapiro.test(anova.ds.sqrt$residuals) #0.0009698
par(mfrow = c(1, 2), oma = c(0, 0, 1.1, 0))
plot(anova.ds, main = "ANOVA", which = 2, ask = F,sub.caption = "")
plot(anova.ds.sqrt, main = "Square Root ANOVA", which = 2, ask = F,sub.caption = "")

```

<font size="1"> Figure No. 9 Normality of Residuals Disease Incidence</font> 

**1.4 Number of stems**

The Duncan and Tukey tests had few differences.Only in year 2016 in cultivar Yuna in the control it can be seen a difference between the groups (Figures 10 and 11). The Shapiro-Wilk normality test was performed for Anova and Square Root Anova because in Anova p was lower than 0.05, that means that the data significantly deviate from a normal distribution (Figure 12).

Taking into account the number of stems, at week 10, we can see that the BZR 336 treatment generated the highest number of stems in both years and in both cultivars in both tests. However, this difference was not significant between the two treatments on cultivar Yuna in 2015, in both tests.

```{r include = FALSE}
#Stems
#Anova Stems
model<-aov(Stems~Treatment+`Weeks after planting`+Year, data=Data.Cultivars.Final)
summary(model)
Data.Cultivars.Final$`Weeks after planting` <- factor(Data.Cultivars.Final$`Weeks after planting`)

#Ducant Test Stems
Y <- W <- Tr <- L <- C <- M <- numeric()

for (i in 2015:2016) { 
  for (k in levels(as.factor(Data.Cultivars.Final$Cultivar))) {
    for(j in levels(Data.Cultivars.Final$`Weeks after planting`)) {
      d1 <- Data.Cultivars.Final[Data.Cultivars.Final$Year==i & 
                                   Data.Cultivars.Final$`Weeks after planting`==j &
                                   Data.Cultivars.Final$Cultivar==k, ]
      model <- aov(Stems~Treatment, data=d1)
      summary(model)
      DT <- duncan.test(model, "Treatment", console = TRUE)
      Y <- c(Y, i,i,i)
      W <- c(W, j,j,j)
      C <- c(C, k,k,k)
      Tr <- c(Tr, row.names(DT$groups))
      L <- c(L, DT$groups$groups)
      M <- c(M, DT$groups$Stems)
    } 
  }
}


LettersST <- data.frame(Year = Y, `Weeks after planting` = factor(W), 
                        Cultivar = C, Treatment = factor(Tr), letra = L, mea = M)

#Tukey Graph Stems

Y <- W <- Tr <- L <- C <- M <- numeric()

for (i in 2015:2016) { 
  for (k in levels(as.factor(Data.Cultivars.Final$Cultivar))) {
    for(j in levels(Data.Cultivars.Final$`Weeks after planting`)) {
      d1z <- Data.Cultivars.Final[Data.Cultivars.Final$Year==i & 
                                    Data.Cultivars.Final$`Weeks after planting`==j &
                                    Data.Cultivars.Final$Cultivar==k, ]
      model <- aov(Stems~Treatment, data=d1z)
      summary(model)
      DT <- HSD.test(model, "Treatment", console = TRUE)
      Y <- c(Y, i,i,i)
      W <- c(W, j,j,j)
      C <- c(C, k,k,k)
      Tr <- c(Tr, row.names(DT$groups))
      L <- c(L, DT$groups$groups)
      M <- c(M, DT$groups$Stems)
    } 
  }
}


LettersSTz <- data.frame(Year = Y, `Weeks after planting` = factor(W), 
                         Cultivar = C, Treatment = factor(Tr), letra = L, mea = M)
```

```{r}
#Duncan Graph Stems 
ggplot(Data.Cultivars.Final, aes(x=`Weeks after planting`, y=Stems, fill=Treatment))+ylab("Number of Stems")+
  geom_boxplot()+ 
  facet_grid(Cultivar~Year)+
  theme_light()+
  geom_text(data = LettersST,
            mapping = aes(x = Weeks.after.planting,
                          y = mea + 1,
                          label = letra),
            position = position_dodge(0.9))+
  labs(caption = "Figure No 10. Results of Duncan Test in the variable Number of Stems") +
  theme(plot.caption.position = "plot",
        plot.caption = element_text(hjust = 0))

#Tukey Graph Stems
ggplot(Data.Cultivars.Final, aes(x=`Weeks after planting`, y=Stems, fill=Treatment))+ylab("Number of Stems")+
  geom_boxplot()+ 
  facet_grid(Cultivar~Year)+
  theme_light()+
  geom_text(data = LettersSTz,
            mapping = aes(x = Weeks.after.planting,
                          y = mea + 1,
                          label = letra),
            position = position_dodge(0.9))+
  labs(caption = "Figure No 11. Results of Tukey Test in the variable Number of Stems") +
  theme(plot.caption.position = "plot",
        plot.caption = element_text(hjust = 0))

#Residuals Stems
anova.st<-aov(Stems~Treatment+`Weeks after planting`+Year, data=Data.Cultivars.Final) #anova 
anova.st.sqrt <- aov(sqrt(Stems)~Treatment+`Weeks after planting`+Year, data=Data.Cultivars.Final) #square root 
shapiro.test(anova.st$residuals) #0.007876 anormal  
shapiro.test(anova.st.sqrt$residuals) #0.002167
par(mfrow = c(1, 2), oma = c(0, 0, 1.1, 0))
plot(anova.ds, main = "ANOVA", which = 2, ask = F,sub.caption = "")
plot(anova.ds.sqrt, main = "Square Root ANOVA", which = 2, ask = F,sub.caption = "")

```

<font size="1"> Figure No. 12 Normality of Residuals Stems</font>


**2) Effect of *B. velezens* strains on the biomass distribution of daughter tubers of two potato cultivars differed by disease resistance to *R. solani* and total yield during 2015 and 2016 growing seasons in Western Siberia.**

**2.1 Biomass distribution**

The Duncan and Tukey tests did not have differences (Figures 13 and 14). The Shapiro-Wilk normality test was performed for Anova and Square Root Anova but this case in Anova p was higher than 0.05, that means that the data has a normal distribution (Figure 15).

Taking into account the biomass, in the medium fraction, we can observe that in both tests and in both years, in cultivar Svitanok kievski, the BZR 517 treatment generated the highest biomass. However, for cultivar Yuna, the difference between treatments and/or control was not significant.


```{r include = FALSE}
#Prueba Biomass
model<-aov(Percentage.Biomass~Treatment+`Fractions`+Year, data=BiomassDistribution)
summary(model)
BiomassDistribution$`Fractions` <- factor(BiomassDistribution$`Fractions`)

#Biomass Duncant Test
Y <- W <- Tr <- L <- C <- M <- numeric()

for (i in 2015:2016) { 
  for (k in levels(as.factor(BiomassDistribution$Cultivar))) {
    for(j in levels(BiomassDistribution$`Fractions`)) {
      d200 <- BiomassDistribution[BiomassDistribution$Year==i & 
                                   BiomassDistribution$`Fractions`==j &
                                   BiomassDistribution$Cultivar==k, ]
      model <- aov(Percentage.Biomass~Treatment, data=d200)
      summary(model)
      DT <- duncan.test(model, "Treatment", console = TRUE)
      Y <- c(Y, i,i,i)
      W <- c(W, j,j,j)
      C <- c(C, k,k,k)
      Tr <- c(Tr, row.names(DT$groups))
      L <- c(L, DT$groups$groups)
      M <- c(M, DT$groups$Percentage.Biomass)
    } 
  }
}

LettersPH1 <- data.frame(Year = Y, `Fractions` = factor(W), 
                         Cultivar = C, Treatment = factor(Tr), letra = L, mea = M)



#Prueba Biomass tukey 
library(agricolae)
Y <- W <- Tr <- L <- C <- M <- numeric()

for (i in 2015:2016) { 
  for (k in levels(as.factor(BiomassDistribution$Cultivar))) {
    for(j in levels(BiomassDistribution$`Fractions`)) {
      t200 <- BiomassDistribution[BiomassDistribution$Year==i & 
                                    BiomassDistribution$`Fractions`==j &
                                    BiomassDistribution$Cultivar==k, ]
      model <- aov(Percentage.Biomass~Treatment, data=t200)
      summary(model)
      DT <- HSD.test(model, "Treatment", console = TRUE)
      Y <- c(Y, i,i,i)
      W <- c(W, j,j,j)
      C <- c(C, k,k,k)
      Tr <- c(Tr, row.names(DT$groups))
      L <- c(L, DT$groups$groups)
      M <- c(M, DT$groups$Percentage.Biomass)
    } 
  }
}
LetterstPH11 <- data.frame(Year = Y, `Fractions` = factor(W), 
                           Cultivar = C, Treatment = factor(Tr), letra = L, mea = M)
```

```{r}
#Biomass Graph Duncan 
ggplot(BiomassDistribution, aes(x=`Fractions`, y=Percentage.Biomass, fill=Treatment))+ylab("Biomass (%)")+
  geom_boxplot()+
  facet_grid(Cultivar~Year)+
  theme_light()+
  geom_text(data = LettersPH1,
            mapping = aes(x = Fractions,
                          y = mea + 8,
                          label = letra),
            position = position_dodge(0.9))+
  labs(caption = "Figure No. 13  Result of Duncan Test on the variable  Percentage of Biomass of potato") +
  theme(plot.caption.position = "plot",
        plot.caption = element_text(hjust = 0))

#Biomass Graph Tukey 
ggplot(BiomassDistribution, aes(x=`Fractions`, y=Percentage.Biomass, fill=Treatment))+ ylab("Biomass (%)")+
  geom_boxplot()+
  facet_grid(Cultivar~Year)+
  theme_light()+
  geom_text(data = LetterstPH11,
            mapping = aes(x = Fractions,
                          y = mea + 8,
                          label = letra),
            position = position_dodge(0.9))+
  labs(caption = "Figure No. 14 Result of Tukey Test in the variable Percentage of Biomass of potato") +
  theme(plot.caption.position = "plot",
        plot.caption = element_text(hjust = 0))

#Residuals Biomass
anova.biomass<-aov(Percentage.Biomass~Treatment+`Fractions`+Year, data=BiomassDistribution) #anova 
anova.biomass.sqrt <- aov(sqrt(Percentage.Biomass)~Treatment+`Fractions`+Year, data=BiomassDistribution) #square root 
shapiro.test(anova.biomass$residuals) #0.06424
shapiro.test(anova.biomass.sqrt$residuals) #0.3364
plot(anova.biomass, main = "ANOVA Percentage of Biomass", which = 2, ask = F,sub.caption = "")
```

<font size="1"> Figure No. 15 Normality of Residuals Percentage of Biomass</font>

**2.2 Yield of Potato**

The Duncan and Tukey tests did not have differences (Figures 16 and 17). The Shapiro-Wilk normality test was performed for Anova with a p-vañue higher than 0.05 both years and cultivars, that means that the data has a normal distribution (Figure 18).

Taking into account the yield of potato, we can observe that in both tests and in both years, in cultivar Svitanok kievski, the BZR 336g treatment generated the highest yield. However, for the cultivar Yuna, the difference between the two treatments was not significant at any time.


```{r include = FALSE}
#Duncan Yield 2015 Svitanok kievskiy
YieldOfPotato$Cultivar <- factor(YieldOfPotato$Cultivar)
Con <- YieldOfPotato$Cultivar == "Svitanok kievskiy" & YieldOfPotato$Year==2015
d1 <- YieldOfPotato[Con,]

model<-aov(Yield.Potato~Treatment,data=d1)
out <- duncan.test(model,"Treatment", 
                   main="Yield of potato")
plot(out,variation="IQR")
out$groups$groups

letter1 <- character()
letter1 <- c(letter1,out$groups$groups)


#Duncan Yield 2015 Yuna
YieldOfPotato$Cultivar <- factor(YieldOfPotato$Cultivar)
Con1 <- YieldOfPotato$Cultivar == "Yuna" & YieldOfPotato$Year==2015
d2 <- YieldOfPotato[Con1,]

model<-aov(Yield.Potato~Treatment,data=d2)
out <- duncan.test(model,"Treatment", 
                   main="Yield of potato")
plot(out,variation="IQR")
out
letter1 <- c(letter1,out$groups$groups)

#Duncan Yield 2016 Svitanok kievskiy
YieldOfPotato$Cultivar <- factor(YieldOfPotato$Cultivar)
Con1234 <- YieldOfPotato$Cultivar == "Svitanok kievskiy" & YieldOfPotato$Year==2016
d1234 <- YieldOfPotato[Con1234,]

model<-aov(Yield.Potato~Treatment,data=d1234)
out <- duncan.test(model,"Treatment", 
                   main="Yield of potato")
plot(out,variation="IQR")
out
letter1 <- c(letter1,out$groups$groups)

#Duncan Yield 2016 Yuna
YieldOfPotato$Cultivar <- factor(YieldOfPotato$Cultivar)
Con12345 <- YieldOfPotato$Cultivar == "Yuna" & YieldOfPotato$Year==2016
d2345 <- YieldOfPotato[Con12345,]

model<-aov(Yield.Potato~Treatment,data=d2345)
out <- duncan.test(model,"Treatment", 
                   main="Yield of potato")
plot(out,variation="IQR")
out
letter1 <- c(letter1,out$groups$groups)

#Tukey Yield 2015 Svitanok kievskiy
YieldOfPotato$Cultivar <- factor(YieldOfPotato$Cultivar)
Con0 <- YieldOfPotato$Cultivar == "Svitanok kievskiy" & YieldOfPotato$Year==2015
t1 <- YieldOfPotato[Con0,]

model<-aov(Yield.Potato~Treatment,data=t1)
out1 <- HSD.test(model,"Treatment", console =TRUE)
out1$groups$groups

lettert <- character()
lettert <- c(lettert,out1$groups$groups)


#Tukey Yield 2015 Yuna
YieldOfPotato$Cultivar <- factor(YieldOfPotato$Cultivar)
Con10 <- YieldOfPotato$Cultivar == "Yuna" & YieldOfPotato$Year==2015
t2 <- YieldOfPotato[Con10,]

model<-aov(Yield.Potato~Treatment,data=t2)
out1 <- HSD.test(model,"Treatment", console =TRUE)
lettert <- c(lettert,out1$groups$groups)


#Tukey Yield 2016 Svitanok kievskiy
YieldOfPotato$Cultivar <- factor(YieldOfPotato$Cultivar)
Cont <- YieldOfPotato$Cultivar == "Svitanok kievskiy" & YieldOfPotato$Year==2016
t3 <- YieldOfPotato[Cont,]

model<-aov(Yield.Potato~Treatment,data=t3)
out1 <- HSD.test(model,"Treatment", console =TRUE)
lettert <- c(lettert,out1$groups$groups)

#Tukey Yield 2016 Yuna
YieldOfPotato$Cultivar <- factor(YieldOfPotato$Cultivar)
Con12345 <- YieldOfPotato$Cultivar == "Yuna" & YieldOfPotato$Year==2016
t4 <- YieldOfPotato[Con12345,]

model<-aov(Yield.Potato~Treatment,data=t4)
out1 <- HSD.test(model,"Treatment", console =TRUE)
lettert <- c(lettert,out1$groups$groups)
```

```{r}
#Duncan Yield graph
library(ggplot2)
library(ggpubr)
yield123 <- ggerrorplot(YieldOfPotato, x = "Cultivar", y = "Yield.Potato",
                        color = "Treatment", palette = "Paired", 
                        error.plot = "pointrange",
                        position = position_dodge(0.5),xlab = ("Cultivars"), ylab = ("Yield of potato (t/ha)"))+
  facet_grid(.~Year)+
  theme_light()
#yield123

#add letterss
dat_text <- data.frame(
  label = letter1,
  Year   = c(rep(2015,6),rep(2016,6)),
  x     = c(0.85, 1,1.2,1.83,2,2.17,0.85, 1,1.2,1.83,2,2.17),
  y     = c(29.3, 23.3,21.6, 27,26,21.8,19.8,17.3,12.3,20.55,20.6,16.1)
)

yield1234 <- yield123 + geom_text(
  data    = dat_text,
  mapping = aes(x = x, y = y, label = label)
)+
  labs(caption = "Figure No. 16 Results of Duncan test Yield of potato") +
  theme(plot.caption.position = "plot",
        plot.caption = element_text(hjust = 0))
yield1234

#Graph Tukey Yield 
library(ggplot2)
library(ggpubr)
yieldtuk <- ggerrorplot(YieldOfPotato, x = "Cultivar", y = "Yield.Potato",
                        color = "Treatment", palette = "Paired", 
                        error.plot = "pointrange",
                        position = position_dodge(0.5),xlab = ("Cultivars"), ylab = ("Yield of potato (t/ha)"))+
  facet_grid(.~Year)+
  theme_light()
#yieldtuk

#add letterss
#c("a", "b", "c","a","a","b","a", "b", "c","a","a","b")
dat_text1 <- data.frame(
  label = lettert,
  Year   = c(rep(2015,6),rep(2016,6)),
  x     = c(0.85, 1,1.2,1.83,2,2.17,0.85, 1,1.2,1.83,2,2.17),
  y     = c(29.3, 23.3,21.6, 27,26,21.8,19.8,17.3,12.3,20.55,20.6,16.1)
)

yieldtukey1 <- yieldtuk + geom_text(
  data    = dat_text1,
  mapping = aes(x = x, y = y, label = label)
)+
  labs(caption = "Figure No. 17 Results of Tukey test Yield of potato") +
  theme(plot.caption.position = "plot",
        plot.caption = element_text(hjust = 0))
yieldtukey1

#Residuals 

#Residuals Yield 2015 Svitanok kievskiy
YieldOfPotato$Cultivar <- factor(YieldOfPotato$Cultivar)
Con <- YieldOfPotato$Cultivar == "Svitanok kievskiy" & YieldOfPotato$Year==2015
d1 <- YieldOfPotato[Con,]
anova.yieldsk<-aov(Yield.Potato~Treatment,data=d1) #anova 
anova.yieldsk.sqrt <- aov(sqrt(Yield.Potato)~Treatment,data=d1) #square root 
shapiro.test(anova.yieldsk$residuals) #0.0553
#shapiro.test(anova.yieldsk.sqrt$residuals) #0.02795

#Residuals Yield 2015 Yuna
YieldOfPotato$Cultivar <- factor(YieldOfPotato$Cultivar)
Con1 <- YieldOfPotato$Cultivar == "Yuna" & YieldOfPotato$Year==2015
d2 <- YieldOfPotato[Con1,]
anova.yieldy<-aov(Yield.Potato~Treatment,data=d2)#anova 
anova.yieldy.sqrt <- aov(sqrt(Yield.Potato)~Treatment,data=d2) #square root
shapiro.test(anova.yieldy$residuals) #0.4242
#shapiro.test(anova.yieldy.sqrt$residuals) #0.3887

#Residuals Yield 2016 Svitanok kievskiy
YieldOfPotato$Cultivar <- factor(YieldOfPotato$Cultivar)
Con1234 <- YieldOfPotato$Cultivar == "Svitanok kievskiy" & YieldOfPotato$Year==2016
d1234 <- YieldOfPotato[Con1234,]
anova.yieldsk6<-aov(Yield.Potato~Treatment,data=d1234)#anova
anova.yieldsk6.sqrt <- aov(sqrt(Yield.Potato)~Treatment,data=d1234)#square root
shapiro.test(anova.yieldsk6$residuals) #0.1674
#shapiro.test(anova.yieldsk6.sqrt$residuals) #0.1481

#Residuals Yield 2016 Yuna
YieldOfPotato$Cultivar <- factor(YieldOfPotato$Cultivar)
Con12345 <- YieldOfPotato$Cultivar == "Yuna" & YieldOfPotato$Year==2016
d2345 <- YieldOfPotato[Con12345,]
anova.yieldy6<-aov(Yield.Potato~Treatment,data=d2345) #anova
anova.yieldy6.sqrt <- aov(sqrt(Yield.Potato)~Treatment,data=d2345) #square root
shapiro.test(anova.yieldy6$residuals) #0.6108
#shapiro.test(anova.yieldy6.sqrt$residuals) #0.6299

#Graph 
par(mfrow = c(2, 2), oma = c(0, 0, 1.1, 0))
p1 <- plot(anova.yieldsk, main = "ANOVA 2015 SK", which = 2, ask = F,sub.caption = "") 
p2 <- plot(anova.yieldy, main = "ANOVA 2015 Y", which = 2, ask = F,sub.caption = "") 
p3 <- plot(anova.yieldsk6, main = "ANOVA 2016 SK", which = 2, ask = F,sub.caption = "") 
p4 <- plot(anova.yieldy6, main = "ANOVA 2016 y", which = 2, ask = F,sub.caption = "")
```

<font size="1"> Figure No. 18 Normality of Residuals Yield of Potato</font>

**3) *In vitro* antifungal activity of *Bacillus strains* against the phytopathogenic fungus *R. solani*, isolated from infected potato tubers**

**3.1 Growth of the diameter of the colonies of *R. solani* under the influence of *B. velenzis* after 3, 5 and 7 days**

Significant differences were observed between the treatments and control over time (Figure 19).


```{r include = FALSE}
##### 1. ANTAGONISTIC ACTIVITY

#### Antagonistic activity was evaluated by measuring the diameter of the fungal 
#### colonies after 3, 5, and 7 days. 

#### All results were expressed and analyzed as the mean.

#### 1.1 Mean and Standard error for Fungal diameter for each treatment.

## Creating a new plot with means and standard errors

#install.packages("dplyr")

library("dplyr")

plot.means <- AntagonisticActivity %>%
  group_by(Days,Trat) %>%
  summarize(n = n(),
            mean = mean(Colony.Diameter),
            sd = sd(Colony.Diameter),
            se = sd / sqrt(n))
```


```{r}
## Graph the fungal diameter growth after 3, 5 and 7 days.

#install.packages("ggplot2")

library(ggplot2)

ggplot(plot.means, aes(x = Days,
                          y = mean,
                          group=Trat,
                          color=Trat)) + guides(col=guide_legend("Treatments"))+
  geom_point(size = 3) +
  geom_line(size = 1) +
  geom_errorbar(aes(ymin=mean-se,
                    ymax=mean+se),
                width = .1)+
  labs(title = "Figure 19. The influence of B. velezensis strains on the growth of R. solani in vitro",
       subtitle = "(Colony diameter)",
       x = "Days",
       y = "Colony diameter",
       color = "Trat") +
  theme_light()
```

## Conclusions 

It can be said that the article in question has very little data to be worked on and therefore the differences between the Tukey and Duncan tests were not significant, which also justifies the authors' choice of use the Duncan test.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


