---
title: "Biological control of important fungal diseases of potato and raspberry by two *Bacillus velezensis* strains"
author: "Lourdes Arellano M, Antonia Fernandes, Juliana Cajiao and Rosario Heribia"
date: "21/12/2021"
output:
  html_document:
    code_folding: hide
---
## Introduction 


## Objectives

Demonstrate the potential effect of *B. velezensis* strains  against *R. solani* as a biological control, and a potential increasement in the plant growth of potato. 

And in this review, the aim of this group is to compare the analysis done in the article between the Duncan and Turkey tests to understand why Duncan was used in the article.


## Hypothesis

*B. velezensis* strains can be used as biological control agents against *R. solani*, and increase significantly the plant growth of potato. Moreover, there is no significant difference between the results obtained through the Duncan and Turkey test.

## Methods


**1. Dowload Data:**
First, dowload the raw data of the paper that is in the link: 
<https://dfzljdn9uc3pi.cloudfront.net/2021/11578/1/peerj-55829-Raw_data.xls>

**Article:**  <https://peerj.com/articles/11578/#>

```{r}
if(!dir.exists("data")) dir.create("data")
download.file("https://dfzljdn9uc3pi.cloudfront.net/2021/11578/1/peerj-55829-Raw_data.xls",
              "data/rawdata.xls")
```

**2. Open and Arrange Data:** For this step, it was used the package "Readxl" to charge the excel file. Each sheet of excel was opened indicating its name and the range required. In some sheets it was use the function "fill" to fill the cells that were empty. Also some names of variables were changed with "rename" and "pivot longer" was used to increase the number of rows or to decrease the number of columns.The potence format was changed. Finally, some new sheets were created why others where joined or deleted in order to have the data arranged. 

```{r include = FALSE}
library(readxl)
readxl::excel_sheets("data/peerj-55829-Raw_data.xls")
dir("data")
#dir("C:\Users\lsare\Desktop\Plant Health\Estadistica\Biologicalcontrol\data\")
library(readxl)
#AntagonisticActivity <- read_excel("data/peerj-55829-Raw_data.xls")
```


```{r include= FALSE}
#Open sheets with ranges 
library
AntagonisticActivity <-
  read_excel("data/peerj-55829-Raw_data.xls",
             sheet = "Antagonistic activity",
             range = "A3:E38")
DiseaseIncidenceAndSeverity <- read_excel("data/peerj-55829-Raw_data.xls",
                                          sheet = "disease incidence and severity",
                                          range = "A4:N40")
MorphometricCharactOfPotato <- read_excel("data/peerj-55829-Raw_data.xls",
                                          sheet = "morphometric charact of potato",)
BiomassDistribution <- read_excel("data/peerj-55829-Raw_data.xls",sheet = "biomass distribution")
YieldOfPotato <- read_excel("data/peerj-55829-Raw_data.xls",
                            sheet = "yield of potato",
                            range = "B3:I15")
#Fill 
library(tidyr)
library(dplyr)
AntagonisticActivity <- AntagonisticActivity %>% fill(everything(), .direction = "down")
BiomassDistribution <- BiomassDistribution %>% fill(everything(), .direction = "down")
DiseaseIncidenceAndSeverity <- DiseaseIncidenceAndSeverity %>% fill(everything(), .direction = "down")
MorphometricCharactOfPotato <- MorphometricCharactOfPotato %>% fill(everything(), .direction = "down")
YieldOfPotato <- YieldOfPotato %>% fill(everything(), .direction = "down")

#Rename 
library(dplyr)
names(AntagonisticActivity)[names(AntagonisticActivity) == 'strain'] <- "Strain"
names(AntagonisticActivity)[names(AntagonisticActivity) == 'CFU/ml'] <- "Concentration"

#Pivot longer 
AntagonisticActivity <- pivot_longer(AntagonisticActivity,cols= c("3d", "5d", "7d"), names_to="Days", values_to="Colony.Diameter")

#change potence
AntagonisticActivity$Concentration
a <- as.numeric(substr(AntagonisticActivity$Concentration, 1, 2))
b <- as.numeric(substr(AntagonisticActivity$Concentration, 3, 3))
a^b
AntagonisticActivity$Concentration <- a^b
rm(a,b)

#Create new column for treatments
AntagonisticActivity$Trat <- paste(AntagonisticActivity$Strain,AntagonisticActivity$Concentration)

#Rename days
library(dplyr)
AntagonisticActivity$Days<-gsub("d","",as.character(AntagonisticActivity$Days))

#Join data frame 
library(dplyr)
Data.Cultivars <- full_join(MorphometricCharactOfPotato, DiseaseIncidenceAndSeverity, by = c("Cultivar","Year", "Treatment","Weeks after planting"))

#Rename 
library(dplyr)
names(Data.Cultivars)[names(Data.Cultivars) == 'Plant height, cm'] <- "1"
names(Data.Cultivars)[names(Data.Cultivars) == '...6.x'] <- "2"
names(Data.Cultivars)[names(Data.Cultivars) == '...7.x'] <- "3"
names(Data.Cultivars)[names(Data.Cultivars) == '...8.x'] <- "4"
names(Data.Cultivars)[names(Data.Cultivars) == '...9.x'] <- "5"

#Pivot longer 
Data.Cultivars <- pivot_longer(Data.Cultivars,cols= c("1", "2", "3","4","5"), names_to="Rep", values_to="Plant.Height")

#Create empty new columns 
Data.Cultivars$Disease.Incidence <- NA
Data.Cultivars$Disease.Severity <- NA
Data.Cultivars$Stems <- NA

#Add data to the columns 
variables <- c("Disease incidence, %","...6.y","...7.y","...8.y","...9.y")
for (i in 1:5) {
  A <- Data.Cultivars$Rep==i
  Data.Cultivars$Disease.Incidence[A] <- Data.Cultivars[[variables[i]]][A]
}

Data.Cultivars$Stems <- NA
variables1 <- c("The number of stems per plant","...11.x","...12.x","...13.x","...14.x")
for (i in 1:5) {
  A1 <- Data.Cultivars$Rep==i
  Data.Cultivars$Stems[A1] <- Data.Cultivars[[variables1[i]]][A1]
}

Data.Cultivars$Disease.Severity <- NA
variables1 <- c("Disease severity, %","...11.y","...12.y","...13.y","...14.y")
for (i in 1:5) {
  A2 <- Data.Cultivars$Rep==i
  Data.Cultivars$Disease.Severity[A2] <- Data.Cultivars[[variables1[i]]][A2]
}


#dataframe with variables that I need 
namesdatacultivar <- c("Cultivar","Treatment","Year","Weeks after planting", "Rep","Plant.Height","Disease.Incidence","Disease.Severity","Stems")           
Data.Cultivars.Final <- Data.Cultivars [which(names(Data.Cultivars) %in% namesdatacultivar)]

#Yieldofpotato 
#Rename 
library(dplyr)
names(YieldOfPotato)[names(YieldOfPotato) == 'Yield, t ha-1'] <- "1"
names(YieldOfPotato)[names(YieldOfPotato) == '...5'] <- "2"
names(YieldOfPotato)[names(YieldOfPotato) == '...6'] <- "3"
names(YieldOfPotato)[names(YieldOfPotato) == '...7'] <- "4"
names(YieldOfPotato)[names(YieldOfPotato) == '...8'] <- "5"
#Pivot longer 
YieldOfPotato <- pivot_longer(YieldOfPotato,cols= c("1", "2", "3","4","5"), names_to="Rep", values_to="Yield.Potato")

#BiomassDistribution
#Rename 
library(dplyr)
names(BiomassDistribution)[names(BiomassDistribution) == 'Biomass distribution by fractions, %'] <- "1"
names(BiomassDistribution)[names(BiomassDistribution) == '...6'] <- "2"
names(BiomassDistribution)[names(BiomassDistribution) == '...7'] <- "3"
names(BiomassDistribution)[names(BiomassDistribution) == '...8'] <- "4"
names(BiomassDistribution)[names(BiomassDistribution) == '...9'] <- "5"
#Pivot longer 
BiomassDistribution <- pivot_longer(BiomassDistribution,cols= c("1", "2", "3","4","5"), names_to="Rep", values_to="Percentage.Biomass")

#create a sheet for medium fraction
medium.fraction <- BiomassDistribution[BiomassDistribution$Fractions=="medium",]

#join sheets of yield and medium fraction
names(YieldOfPotato)[names(YieldOfPotato) == 'Bacillus strain'] <- "Treatment"
YieldOfPotato <- full_join(YieldOfPotato,medium.fraction)
YieldOfPotato <- YieldOfPotato[,-which(names(YieldOfPotato)=="Fractions")]

#remove sheets
ls()
rm("Data.Cultivars","DiseaseIncidenceAndSeverity","medium.fraction","MorphometricCharactOfPotato")

```


**3. Do analysis:**
Finally with the arrangement of the sheets of excel, there were obtained 4 main data sets. The first one was Antagonistic Activity, then Data Cultivars that include the variables of Percentage of Disease Incidence, Percentage of Disease Severity, Plant height and number of stems per plant.Then, the data set of Biomass Distribution and finally the data set of Yield of Potato. 

With these data sets were performed Tukey HSD Test, Duncan test and Normality Test of Residuals with shapiro Test. All results were expressed in graphs using the packages of "ggplot" and "ggpubr". The results were analyzed and compared with the results of the article. It is important to say that in the article, the authors have only performed Duncan Test.

**Description of data sets:**

**1) Effect of *B. velezensis* strains on disease incidence and severity and morphometric parameters caused by *R. solani* on two potato cultivars differed by disease resistance during 2015 and 2016 growing seasons in Western Siberia**


Cultivars:

1. Svitanok kievskiy 

2. Yuna


Years:

* 2015

* 2016


Treatments:

1. Control

2. BZR 336g 

3. BZR 517


Weeks after planting:

* 4

* 6

* 10


Variables: (5 rep/treatment)

1. Disease incidence (%) 

2. Disease severity (%)

3. Plant height (cm) 

4. Number of stems per plant


&nbsp;

**2) Effect of *B. velezens* strains on the biomass distribution of daughter tubers of two potato cultivars differed by disease resistance to *R. solani* and total yield during 2015 and 2016 growing seasons in Western Siberia.**


Cultivars:

1. Svitanok kievskiy 

2. Yuna


Years: 

* 2015

* 2016


Treatments:

1. Control 

2. BZR 336g

3. BZR 517


Fractions:

* Small

* Medium

* Large 


Variables: (5 rep/treatment)

1. Biomass distribution 

**3) Antifungal activity of B. velezensis against *R solani* in vitro.** 

Strains:

1. *Bacillus velezensis* strain BZR 336 g

2. *Bacillus velezensis* strain BZR 517

3. Control 


Treatments:
 
1. 10^4 CFU/mL
  
2. 10^5 CFU/mL
  
3. 10^6 CFU/mL
  
4. Control.

###### CFU: Colony-Forming Unit


Days:

* 3

* 5

* 7 

Variables: 

* Colony Diameter (cm)


&nbsp;

## Results and Discussion 


&nbsp;

**1) Effect of *B. velezensis* strains on disease incidence and severity and morphometric parameters caused by *R. solani* on two potato cultivars differed by disease resistance during 2015 and 2016 growing seasons in Western Siberia**


&nbsp;
**1.1 Plant Height**

The Duncan and Tukey tests had few differences.Only in year 2016, week 6 in cultivar Svitanok kievski it can be seen a difference in the group of BZR 517 (Figures 1 and 2). The Shapiro-Wilk normality test was performed for Anova and Square Root Anova because in Anova p was lower than 0.05, that means that the data significantly deviate from a normal distribution (Figure 3).

```{r include = FALSE}
#Duncan Analysis Data.Cultivars.Final
library(agricolae)
library(ggplot2)
library(ggpubr)

# Plant Height

#Anova Plant Height
model<-aov(Plant.Height~Treatment+`Weeks after planting`+Year, data=Data.Cultivars.Final)
summary(model)
Data.Cultivars.Final$`Weeks after planting` <- factor(Data.Cultivars.Final$`Weeks after planting`)

## Plant Height Duncan 

Y <- W <- Tr <- L <- C <- M <- numeric()

for (i in 2015:2016) { 
  for (k in levels(as.factor(Data.Cultivars.Final$Cultivar))) {
    for(j in levels(Data.Cultivars.Final$`Weeks after planting`)) {
      d1 <- Data.Cultivars.Final[Data.Cultivars.Final$Year==i & 
                                   Data.Cultivars.Final$`Weeks after planting`==j &
                                   Data.Cultivars.Final$Cultivar==k, ]
      model <- aov(Plant.Height~Treatment, data=d1)
      summary(model)
      DT <- duncan.test(model, "Treatment", console = TRUE)
      Y <- c(Y, i,i,i)
      W <- c(W, j,j,j)
      C <- c(C, k,k,k)
      Tr <- c(Tr, row.names(DT$groups))
      L <- c(L, DT$groups$groups)
      M <- c(M, DT$groups$Plant.Height)
    } 
  }
}


LettersPH <- data.frame(Year = Y, `Weeks after planting` = factor(W), 
                        Cultivar = C, Treatment = factor(Tr), letra = L, mea = M)

## Plant Height Tukey 

Y <- W <- Tr <- L <- C <- M <- numeric()

for (i in 2015:2016) { 
  for (k in levels(as.factor(Data.Cultivars.Final$Cultivar))) {
    for(j in levels(Data.Cultivars.Final$`Weeks after planting`)) {
      d1z <- Data.Cultivars.Final[Data.Cultivars.Final$Year==i & 
                                    Data.Cultivars.Final$`Weeks after planting`==j &
                                    Data.Cultivars.Final$Cultivar==k, ]
      model <- aov(Plant.Height~Treatment, data=d1z)
      summary(model)
      DT <- HSD.test(model, "Treatment", console = TRUE)
      Y <- c(Y, i,i,i)
      W <- c(W, j,j,j)
      C <- c(C, k,k,k)
      Tr <- c(Tr, row.names(DT$groups))
      L <- c(L, DT$groups$groups)
      M <- c(M, DT$groups$Plant.Height)
    } 
  }
}

LettersPHz <- data.frame(Year = Y, `Weeks after planting` = factor(W), 
                         Cultivar = C, Treatment = factor(Tr), letra = L, mea = M)

```

```{r}
library(agricolae)
library(ggplot2)
library(ggpubr)
#Plant Height Duncan Graph

ggplot(Data.Cultivars.Final, aes(x=`Weeks after planting`, y=Plant.Height, fill=Treatment))+ylab("Plant Height (cm)")+
  geom_boxplot()+
  facet_grid(Cultivar~Year)+
  theme_light()+
  geom_text(data = LettersPH,
            mapping = aes(x = Weeks.after.planting,
                          y = mea + 5,
                          label = letra),
            position = position_dodge(0.9))+
  labs(caption = "Figure No 1. Results of Duncan Test in the variable of Plant Height on potato (cm)") +
  theme(plot.caption.position = "plot",
        plot.caption = element_text(hjust = 0))
```

```{r}
#Plant Height Tukey Graph

ggplot(Data.Cultivars.Final, aes(x=`Weeks after planting`, y=Plant.Height, fill=Treatment))+ylab("Plant Height (cm)")+
  geom_boxplot()+
  facet_grid(Cultivar~Year)+
  theme_light()+
  geom_text(data = LettersPHz,
            mapping = aes(x = Weeks.after.planting,
                          y = mea + 5,
                          label = letra),
            position = position_dodge(0.9))+
  labs(caption = "Figure No 2. Results of Tukey Test in the variable of Plant Height on potato (cm)") +
  theme(plot.caption.position = "plot",
        plot.caption = element_text(hjust = 0))

#Residuals Plant Height 
anova.ph<-aov(Plant.Height~Treatment+`Weeks after planting`+Year, data=Data.Cultivars.Final) #anova 
anova.ph.sqrt <- aov(sqrt(Plant.Height)~Treatment+`Weeks after planting`+Year, data=Data.Cultivars.Final) #square root 
shapiro.test(anova.ph$residuals) #0.0329 anormal 
shapiro.test(anova.ph.sqrt$residuals) #0.02933
par(mfrow = c(1, 2), oma = c(0, 0, 1.1, 0))
plot(anova.ph, main = "ANOVA", which = 2, ask = F,sub.caption = "")
plot(anova.ph.sqrt, main = "Square Root ANOVA", which = 2, ask = F,sub.caption = "")
```

<font size="1.5"> Figure No 3. Normality of Residuals Plant Height.</font> 



&nbsp;

**1.2 Disease Incidence** 

The Duncan and Tukey tests had few differences.Only in year 2015 in cultivar Svitanok kievski it can be seen a difference between the groups (Figures 4 and 5). The Shapiro-Wilk normality test was performed for Anova and Square Root Anova because in Anova p was lower than 0.05, that means that the data significantly deviate from a normal distribution (Figure 6). 

```{r include = FALSE}
#Disease Incidence
#Anova Disease Incidence 
model<-aov(Disease.Incidence~Treatment+`Weeks after planting`+Year, data=Data.Cultivars.Final)
summary(model)
Data.Cultivars.Final$`Weeks after planting` <- factor(Data.Cultivars.Final$`Weeks after planting`)


#Duncan Test Disease Incidence

Y <- W <- Tr <- L <- C <- M <- numeric()

for (i in 2015:2016) { 
  for (k in levels(as.factor(Data.Cultivars.Final$Cultivar))) {
    for(j in levels(Data.Cultivars.Final$`Weeks after planting`)) {
      d1 <- Data.Cultivars.Final[Data.Cultivars.Final$Year==i & 
                                   Data.Cultivars.Final$`Weeks after planting`==j &
                                   Data.Cultivars.Final$Cultivar==k, ]
      model <- aov(Disease.Incidence~Treatment, data=d1)
      summary(model)
      DT <- duncan.test(model, "Treatment", console = TRUE)
      Y <- c(Y, i,i,i)
      W <- c(W, j,j,j)
      C <- c(C, k,k,k)
      Tr <- c(Tr, row.names(DT$groups))
      L <- c(L, DT$groups$groups)
      M <- c(M, DT$groups$Disease.Incidence)
    } 
  }
}


LettersDI <- data.frame(Year = Y, `Weeks after planting` = factor(W), 
                        Cultivar = C, Treatment = factor(Tr), letra = L, mea = M)

#Tukey Disease Incidence

Y <- W <- Tr <- L <- C <- M <- numeric()

for (i in 2015:2016) { 
  for (k in levels(as.factor(Data.Cultivars.Final$Cultivar))) {
    for(j in levels(Data.Cultivars.Final$`Weeks after planting`)) {
      d1z <- Data.Cultivars.Final[Data.Cultivars.Final$Year==i & 
                                    Data.Cultivars.Final$`Weeks after planting`==j &
                                    Data.Cultivars.Final$Cultivar==k, ]
      model <- aov(Disease.Incidence~Treatment, data=d1z)
      summary(model)
      DT <- HSD.test(model, "Treatment", console = TRUE)
      Y <- c(Y, i,i,i)
      W <- c(W, j,j,j)
      C <- c(C, k,k,k)
      Tr <- c(Tr, row.names(DT$groups))
      L <- c(L, DT$groups$groups)
      M <- c(M, DT$groups$Disease.Incidence)
    } 
  }
}


LettersDIz <- data.frame(Year = Y, `Weeks after planting` = factor(W), 
                         Cultivar = C, Treatment = factor(Tr), letra = L, mea = M)
```

```{r}
#Duncan Graph Disease Incidence
ggplot(Data.Cultivars.Final, aes(x=`Weeks after planting`, y=Disease.Incidence, fill=Treatment))+ylab("Disease Incidence(%)")+
  geom_boxplot()+
  facet_grid(Cultivar~Year)+
  theme_light()+
  geom_text(data = LettersDI,
            mapping = aes(x = Weeks.after.planting,
                          y = mea + 8,
                          label = letra),
            position = position_dodge(0.9))+
  labs(caption = "Figure No 4. Results of Duncan Test in the variable of Disease Incidence (%)") +
  theme(plot.caption.position = "plot",
        plot.caption = element_text(hjust = 0))

#Tukey Graph Disease Incidence
ggplot(Data.Cultivars.Final, aes(x=`Weeks after planting`, y=Disease.Incidence, fill=Treatment))+ylab("Disease Incidence(%)")+
  geom_boxplot()+
  facet_grid(Cultivar~Year)+
  theme_light()+
  geom_text(data = LettersDIz,
            mapping = aes(x = Weeks.after.planting,
                          y = mea + 8,
                          label = letra),
            position = position_dodge(0.9))+
  labs(caption = "Figure No 5. Results of Tukey Test in the variable of Disease Incidence (%)") +
  theme(plot.caption.position = "plot",
        plot.caption = element_text(hjust = 0))
#Residuals Disease Incidence
anova.di<-aov(Disease.Incidence~Treatment+`Weeks after planting`+Year, data=Data.Cultivars.Final) #anova 
anova.di.sqrt <- aov(sqrt(Disease.Incidence)~Treatment+`Weeks after planting`+Year, data=Data.Cultivars.Final) #square root 
shapiro.test(anova.di$residuals) #0.03132 anormal 
shapiro.test(anova.di.sqrt$residuals) #0.001165
par(mfrow = c(1, 2), oma = c(0, 0, 1.1, 0))
plot(anova.di, main = "ANOVA", which = 2, ask = F,sub.caption = "")
plot(anova.di.sqrt, main = "Square Root ANOVA", which = 2, ask = F,sub.caption = "")
```

<font size="1"> Figure No. 6 Normality of Residuals Disease Incidence</font> 


**1.3 Disease Severity**

The Duncan and Tukey tests did not have differences (Figures 7 and 8). The Shapiro-Wilk normality test was performed for Anova and Square Root Anova because in Anova p was lower than 0.05, that means that the data significantly deviate from a normal distribution (Figure 9).

```{r include = FALSE}
#Disease Severity

#Anova Disease Severity
model<-aov(Disease.Severity~Treatment+`Weeks after planting`+Year, data=Data.Cultivars.Final)
summary(model)
Data.Cultivars.Final$`Weeks after planting` <- factor(Data.Cultivars.Final$`Weeks after planting`)
#Disease Severity Duncan Test

Y <- W <- Tr <- L <- C <- M <- numeric()

for (i in 2015:2016) { 
  for (k in levels(as.factor(Data.Cultivars.Final$Cultivar))) {
    for(j in levels(Data.Cultivars.Final$`Weeks after planting`)) {
      d1 <- Data.Cultivars.Final[Data.Cultivars.Final$Year==i & 
                                   Data.Cultivars.Final$`Weeks after planting`==j &
                                   Data.Cultivars.Final$Cultivar==k, ]
      model <- aov(Disease.Severity~Treatment, data=d1)
      summary(model)
      DT <- duncan.test(model, "Treatment", console = TRUE)
      Y <- c(Y, i,i,i)
      W <- c(W, j,j,j)
      C <- c(C, k,k,k)
      Tr <- c(Tr, row.names(DT$groups))
      L <- c(L, DT$groups$groups)
      M <- c(M, DT$groups$Disease.Severity)
    } 
  }
}


LettersDS <- data.frame(Year = Y, `Weeks after planting` = factor(W), 
                        Cultivar = C, Treatment = factor(Tr), letra = L, mea = M)

#Tukey Test Disease Severity

Y <- W <- Tr <- L <- C <- M <- numeric()

for (i in 2015:2016) { 
  for (k in levels(as.factor(Data.Cultivars.Final$Cultivar))) {
    for(j in levels(Data.Cultivars.Final$`Weeks after planting`)) {
      d1z <- Data.Cultivars.Final[Data.Cultivars.Final$Year==i & 
                                    Data.Cultivars.Final$`Weeks after planting`==j &
                                    Data.Cultivars.Final$Cultivar==k, ]
      model <- aov(Disease.Severity~Treatment, data=d1z)
      summary(model)
      DT <- HSD.test(model, "Treatment", console = TRUE)
      Y <- c(Y, i,i,i)
      W <- c(W, j,j,j)
      C <- c(C, k,k,k)
      Tr <- c(Tr, row.names(DT$groups))
      L <- c(L, DT$groups$groups)
      M <- c(M, DT$groups$Disease.Severity)
    } 
  }
}


LettersDSz <- data.frame(Year = Y, `Weeks after planting` = factor(W), 
                         Cultivar = C, Treatment = factor(Tr), letra = L, mea = M)
```

```{r}
#Duncan Graph Disease Severity
ggplot(Data.Cultivars.Final, aes(x=`Weeks after planting`, y=Disease.Severity, fill=Treatment))+ylab("Disease Severity (%)")+
  geom_boxplot()+
  facet_grid(Cultivar~Year)+
  theme_light()+
  geom_text(data = LettersDS,
            mapping = aes(x = Weeks.after.planting,
                          y = mea + 5,
                          label = letra),
            position = position_dodge(0.9))+
  labs(caption = "Figure No 7. Results of Duncan Test in the variable of Disease Severity (%)") +
  theme(plot.caption.position = "plot",
        plot.caption = element_text(hjust = 0))

#Tukey Graph Disease Severity

ggplot(Data.Cultivars.Final, aes(x=`Weeks after planting`, y=Disease.Severity, fill=Treatment))+ylab("Disease Severity(%)")+
  geom_boxplot()+
  facet_grid(Cultivar~Year)+
  theme_light()+ 
  geom_text(data = LettersDSz,
            mapping = aes(x = Weeks.after.planting,
                          y = mea + 7,
                          label = letra),
            position = position_dodge(0.9))+
  labs(caption = "Figure No 8. Results of Tukey Test in the variable of Disease Severity (%)") +
  theme(plot.caption.position = "plot",
        plot.caption = element_text(hjust = 0))
#Residuals Disease Severity
anova.ds<-aov(Disease.Severity~Treatment+`Weeks after planting`+Year, data=Data.Cultivars.Final) #anova 
anova.ds.sqrt <- aov(sqrt(Disease.Severity)~Treatment+`Weeks after planting`+Year, data=Data.Cultivars.Final) #square root 
shapiro.test(anova.ds$residuals) #0.002356 anormal  
shapiro.test(anova.ds.sqrt$residuals) #0.0009698
par(mfrow = c(1, 2), oma = c(0, 0, 1.1, 0))
plot(anova.ds, main = "ANOVA", which = 2, ask = F,sub.caption = "")
plot(anova.ds.sqrt, main = "Square Root ANOVA", which = 2, ask = F,sub.caption = "")

```

<font size="1"> Figure No. 9 Normality of Residuals Disease Incidence</font> 

**1.4 Number of stems**

The Duncan and Tukey tests had few differences.Only in year 2016 in cultivar Yuna in the control it can be seen a difference between the groups (Figures 10 and 11). The Shapiro-Wilk normality test was performed for Anova and Square Root Anova because in Anova p was lower than 0.05, that means that the data significantly deviate from a normal distribution (Figure 12).

```{r include = FALSE}
#Stems
#Anova Stems
model<-aov(Stems~Treatment+`Weeks after planting`+Year, data=Data.Cultivars.Final)
summary(model)
Data.Cultivars.Final$`Weeks after planting` <- factor(Data.Cultivars.Final$`Weeks after planting`)

#Ducant Test Stems
Y <- W <- Tr <- L <- C <- M <- numeric()

for (i in 2015:2016) { 
  for (k in levels(as.factor(Data.Cultivars.Final$Cultivar))) {
    for(j in levels(Data.Cultivars.Final$`Weeks after planting`)) {
      d1 <- Data.Cultivars.Final[Data.Cultivars.Final$Year==i & 
                                   Data.Cultivars.Final$`Weeks after planting`==j &
                                   Data.Cultivars.Final$Cultivar==k, ]
      model <- aov(Stems~Treatment, data=d1)
      summary(model)
      DT <- duncan.test(model, "Treatment", console = TRUE)
      Y <- c(Y, i,i,i)
      W <- c(W, j,j,j)
      C <- c(C, k,k,k)
      Tr <- c(Tr, row.names(DT$groups))
      L <- c(L, DT$groups$groups)
      M <- c(M, DT$groups$Stems)
    } 
  }
}


LettersST <- data.frame(Year = Y, `Weeks after planting` = factor(W), 
                        Cultivar = C, Treatment = factor(Tr), letra = L, mea = M)

#Tukey Graph Stems

Y <- W <- Tr <- L <- C <- M <- numeric()

for (i in 2015:2016) { 
  for (k in levels(as.factor(Data.Cultivars.Final$Cultivar))) {
    for(j in levels(Data.Cultivars.Final$`Weeks after planting`)) {
      d1z <- Data.Cultivars.Final[Data.Cultivars.Final$Year==i & 
                                    Data.Cultivars.Final$`Weeks after planting`==j &
                                    Data.Cultivars.Final$Cultivar==k, ]
      model <- aov(Stems~Treatment, data=d1z)
      summary(model)
      DT <- HSD.test(model, "Treatment", console = TRUE)
      Y <- c(Y, i,i,i)
      W <- c(W, j,j,j)
      C <- c(C, k,k,k)
      Tr <- c(Tr, row.names(DT$groups))
      L <- c(L, DT$groups$groups)
      M <- c(M, DT$groups$Stems)
    } 
  }
}


LettersSTz <- data.frame(Year = Y, `Weeks after planting` = factor(W), 
                         Cultivar = C, Treatment = factor(Tr), letra = L, mea = M)
```

```{r}
#Duncan Graph Stems 
ggplot(Data.Cultivars.Final, aes(x=`Weeks after planting`, y=Stems, fill=Treatment))+ylab("Number of Stems")+
  geom_boxplot()+ 
  facet_grid(Cultivar~Year)+
  theme_light()+
  geom_text(data = LettersST,
            mapping = aes(x = Weeks.after.planting,
                          y = mea + 1,
                          label = letra),
            position = position_dodge(0.9))+
  labs(caption = "Figure No 10. Results of Duncan Test in the variable Number of Stems") +
  theme(plot.caption.position = "plot",
        plot.caption = element_text(hjust = 0))

#Tukey Graph Stems
ggplot(Data.Cultivars.Final, aes(x=`Weeks after planting`, y=Stems, fill=Treatment))+ylab("Number of Stems")+
  geom_boxplot()+ 
  facet_grid(Cultivar~Year)+
  theme_light()+
  geom_text(data = LettersSTz,
            mapping = aes(x = Weeks.after.planting,
                          y = mea + 1,
                          label = letra),
            position = position_dodge(0.9))+
  labs(caption = "Figure No 11. Results of Tukey Test in the variable Number of Stems") +
  theme(plot.caption.position = "plot",
        plot.caption = element_text(hjust = 0))

#Residuals Stems
anova.st<-aov(Stems~Treatment+`Weeks after planting`+Year, data=Data.Cultivars.Final) #anova 
anova.st.sqrt <- aov(sqrt(Stems)~Treatment+`Weeks after planting`+Year, data=Data.Cultivars.Final) #square root 
shapiro.test(anova.st$residuals) #0.007876 anormal  
shapiro.test(anova.st.sqrt$residuals) #0.002167
par(mfrow = c(1, 2), oma = c(0, 0, 1.1, 0))
plot(anova.ds, main = "ANOVA", which = 2, ask = F,sub.caption = "")
plot(anova.ds.sqrt, main = "Square Root ANOVA", which = 2, ask = F,sub.caption = "")

```

<font size="1"> Figure No. 12 Normality of Residuals Stems</font>


**2) Effect of *B. velezens* strains on the biomass distribution of daughter tubers of two potato cultivars differed by disease resistance to *R. solani* and total yield during 2015 and 2016 growing seasons in Western Siberia.**

**2.1 Biomass distribution**

The Duncan and Tukey tests did not have differences (Figures 13 and 14). The Shapiro-Wilk normality test was performed for Anova and Square Root Anova but this case in Anova p was higher than 0.05, that means that the data has a normal distribution (Figure 15).


```{r include = FALSE}
#Prueba Biomass
model<-aov(Percentage.Biomass~Treatment+`Fractions`+Year, data=BiomassDistribution)
summary(model)
BiomassDistribution$`Fractions` <- factor(BiomassDistribution$`Fractions`)

#Biomass Duncant Test
Y <- W <- Tr <- L <- C <- M <- numeric()

for (i in 2015:2016) { 
  for (k in levels(as.factor(BiomassDistribution$Cultivar))) {
    for(j in levels(BiomassDistribution$`Fractions`)) {
      d200 <- BiomassDistribution[BiomassDistribution$Year==i & 
                                   BiomassDistribution$`Fractions`==j &
                                   BiomassDistribution$Cultivar==k, ]
      model <- aov(Percentage.Biomass~Treatment, data=d200)
      summary(model)
      DT <- duncan.test(model, "Treatment", console = TRUE)
      Y <- c(Y, i,i,i)
      W <- c(W, j,j,j)
      C <- c(C, k,k,k)
      Tr <- c(Tr, row.names(DT$groups))
      L <- c(L, DT$groups$groups)
      M <- c(M, DT$groups$Percentage.Biomass)
    } 
  }
}

LettersPH1 <- data.frame(Year = Y, `Fractions` = factor(W), 
                         Cultivar = C, Treatment = factor(Tr), letra = L, mea = M)



#Prueba Biomass tukey 
library(agricolae)
Y <- W <- Tr <- L <- C <- M <- numeric()

for (i in 2015:2016) { 
  for (k in levels(as.factor(BiomassDistribution$Cultivar))) {
    for(j in levels(BiomassDistribution$`Fractions`)) {
      t200 <- BiomassDistribution[BiomassDistribution$Year==i & 
                                    BiomassDistribution$`Fractions`==j &
                                    BiomassDistribution$Cultivar==k, ]
      model <- aov(Percentage.Biomass~Treatment, data=t200)
      summary(model)
      DT <- HSD.test(model, "Treatment", console = TRUE)
      Y <- c(Y, i,i,i)
      W <- c(W, j,j,j)
      C <- c(C, k,k,k)
      Tr <- c(Tr, row.names(DT$groups))
      L <- c(L, DT$groups$groups)
      M <- c(M, DT$groups$Percentage.Biomass)
    } 
  }
}
LetterstPH11 <- data.frame(Year = Y, `Fractions` = factor(W), 
                           Cultivar = C, Treatment = factor(Tr), letra = L, mea = M)
```

```{r}
#Biomass Graph Duncan 
ggplot(BiomassDistribution, aes(x=`Fractions`, y=Percentage.Biomass, fill=Treatment))+ylab("Biomass (%)")+
  geom_boxplot()+
  facet_grid(Cultivar~Year)+
  theme_light()+
  geom_text(data = LettersPH1,
            mapping = aes(x = Fractions,
                          y = mea + 8,
                          label = letra),
            position = position_dodge(0.9))+
  labs(caption = "Figure No. 13  Result of Duncan Test on the variable  Percentage of Biomass of potato") +
  theme(plot.caption.position = "plot",
        plot.caption = element_text(hjust = 0))

#Biomass Graph Tukey 
ggplot(BiomassDistribution, aes(x=`Fractions`, y=Percentage.Biomass, fill=Treatment))+ ylab("Biomass (%)")+
  geom_boxplot()+
  facet_grid(Cultivar~Year)+
  theme_light()+
  geom_text(data = LetterstPH11,
            mapping = aes(x = Fractions,
                          y = mea + 8,
                          label = letra),
            position = position_dodge(0.9))+
  labs(caption = "Figure No. 14 Result of Tukey Test in the variable Percentage of Biomass of potato") +
  theme(plot.caption.position = "plot",
        plot.caption = element_text(hjust = 0))

#Residuals Biomass
anova.biomass<-aov(Percentage.Biomass~Treatment+`Fractions`+Year, data=BiomassDistribution) #anova 
anova.biomass.sqrt <- aov(sqrt(Percentage.Biomass)~Treatment+`Fractions`+Year, data=BiomassDistribution) #square root 
shapiro.test(anova.biomass$residuals) #0.06424
shapiro.test(anova.biomass.sqrt$residuals) #0.3364
plot(anova.biomass, main = "ANOVA Percentage of Biomass", which = 2, ask = F,sub.caption = "")
```

<font size="1"> Figure No. 15 Normality of Residuals Percentage of Biomass</font>

**2.2 Yield of Potato**

The Duncan and Tukey tests did not have differences (Figures 16 and 17). The Shapiro-Wilk normality test was performed for Anova with a p-vañue higher than 0.05 both years and cultivars, that means that the data has a normal distribution (Figure 18).

```{r include = FALSE}
#Duncan Yield 2015 Svitanok kievskiy
YieldOfPotato$Cultivar <- factor(YieldOfPotato$Cultivar)
Con <- YieldOfPotato$Cultivar == "Svitanok kievskiy" & YieldOfPotato$Year==2015
d1 <- YieldOfPotato[Con,]

model<-aov(Yield.Potato~Treatment,data=d1)
out <- duncan.test(model,"Treatment", 
                   main="Yield of potato")
plot(out,variation="IQR")
out$groups$groups

letter1 <- character()
letter1 <- c(letter1,out$groups$groups)


#Duncan Yield 2015 Yuna
YieldOfPotato$Cultivar <- factor(YieldOfPotato$Cultivar)
Con1 <- YieldOfPotato$Cultivar == "Yuna" & YieldOfPotato$Year==2015
d2 <- YieldOfPotato[Con1,]

model<-aov(Yield.Potato~Treatment,data=d2)
out <- duncan.test(model,"Treatment", 
                   main="Yield of potato")
plot(out,variation="IQR")
out
letter1 <- c(letter1,out$groups$groups)

#Duncan Yield 2016 Svitanok kievskiy
YieldOfPotato$Cultivar <- factor(YieldOfPotato$Cultivar)
Con1234 <- YieldOfPotato$Cultivar == "Svitanok kievskiy" & YieldOfPotato$Year==2016
d1234 <- YieldOfPotato[Con1234,]

model<-aov(Yield.Potato~Treatment,data=d1234)
out <- duncan.test(model,"Treatment", 
                   main="Yield of potato")
plot(out,variation="IQR")
out
letter1 <- c(letter1,out$groups$groups)

#Duncan Yield 2016 Yuna
YieldOfPotato$Cultivar <- factor(YieldOfPotato$Cultivar)
Con12345 <- YieldOfPotato$Cultivar == "Yuna" & YieldOfPotato$Year==2016
d2345 <- YieldOfPotato[Con12345,]

model<-aov(Yield.Potato~Treatment,data=d2345)
out <- duncan.test(model,"Treatment", 
                   main="Yield of potato")
plot(out,variation="IQR")
out
letter1 <- c(letter1,out$groups$groups)

#Tukey Yield 2015 Svitanok kievskiy
YieldOfPotato$Cultivar <- factor(YieldOfPotato$Cultivar)
Con0 <- YieldOfPotato$Cultivar == "Svitanok kievskiy" & YieldOfPotato$Year==2015
t1 <- YieldOfPotato[Con0,]

model<-aov(Yield.Potato~Treatment,data=t1)
out1 <- HSD.test(model,"Treatment", console =TRUE)
out1$groups$groups

lettert <- character()
lettert <- c(lettert,out1$groups$groups)


#Tukey Yield 2015 Yuna
YieldOfPotato$Cultivar <- factor(YieldOfPotato$Cultivar)
Con10 <- YieldOfPotato$Cultivar == "Yuna" & YieldOfPotato$Year==2015
t2 <- YieldOfPotato[Con10,]

model<-aov(Yield.Potato~Treatment,data=t2)
out1 <- HSD.test(model,"Treatment", console =TRUE)
lettert <- c(lettert,out1$groups$groups)


#Tukey Yield 2016 Svitanok kievskiy
YieldOfPotato$Cultivar <- factor(YieldOfPotato$Cultivar)
Cont <- YieldOfPotato$Cultivar == "Svitanok kievskiy" & YieldOfPotato$Year==2016
t3 <- YieldOfPotato[Cont,]

model<-aov(Yield.Potato~Treatment,data=t3)
out1 <- HSD.test(model,"Treatment", console =TRUE)
lettert <- c(lettert,out1$groups$groups)

#Tukey Yield 2016 Yuna
YieldOfPotato$Cultivar <- factor(YieldOfPotato$Cultivar)
Con12345 <- YieldOfPotato$Cultivar == "Yuna" & YieldOfPotato$Year==2016
t4 <- YieldOfPotato[Con12345,]

model<-aov(Yield.Potato~Treatment,data=t4)
out1 <- HSD.test(model,"Treatment", console =TRUE)
lettert <- c(lettert,out1$groups$groups)
```

```{r}
#Duncan Yield graph
library(ggplot2)
library(ggpubr)
yield123 <- ggerrorplot(YieldOfPotato, x = "Cultivar", y = "Yield.Potato",
                        color = "Treatment", palette = "Paired", 
                        error.plot = "pointrange",
                        position = position_dodge(0.5),xlab = ("Cultivars"), ylab = ("Yield of potato (t/ha)"))+
  facet_grid(.~Year)+
  theme_light()
#yield123

#add letterss
dat_text <- data.frame(
  label = letter1,
  Year   = c(rep(2015,6),rep(2016,6)),
  x     = c(0.85, 1,1.2,1.83,2,2.17,0.85, 1,1.2,1.83,2,2.17),
  y     = c(29.3, 23.3,21.6, 27,26,21.8,19.8,17.3,12.3,20.55,20.6,16.1)
)

yield1234 <- yield123 + geom_text(
  data    = dat_text,
  mapping = aes(x = x, y = y, label = label)
)+
  labs(caption = "Figure No. 16 Results of Duncan test Yield of potato") +
  theme(plot.caption.position = "plot",
        plot.caption = element_text(hjust = 0))
yield1234

#Graph Tukey Yield 
library(ggplot2)
library(ggpubr)
yieldtuk <- ggerrorplot(YieldOfPotato, x = "Cultivar", y = "Yield.Potato",
                        color = "Treatment", palette = "Paired", 
                        error.plot = "pointrange",
                        position = position_dodge(0.5),xlab = ("Cultivars"), ylab = ("Yield of potato (t/ha)"))+
  facet_grid(.~Year)+
  theme_light()
#yieldtuk

#add letterss
#c("a", "b", "c","a","a","b","a", "b", "c","a","a","b")
dat_text1 <- data.frame(
  label = lettert,
  Year   = c(rep(2015,6),rep(2016,6)),
  x     = c(0.85, 1,1.2,1.83,2,2.17,0.85, 1,1.2,1.83,2,2.17),
  y     = c(29.3, 23.3,21.6, 27,26,21.8,19.8,17.3,12.3,20.55,20.6,16.1)
)

yieldtukey1 <- yieldtuk + geom_text(
  data    = dat_text1,
  mapping = aes(x = x, y = y, label = label)
)+
  labs(caption = "Figure No. 17 Results of Tukey test Yield of potato") +
  theme(plot.caption.position = "plot",
        plot.caption = element_text(hjust = 0))
yieldtukey1

#Residuals 

#Residuals Yield 2015 Svitanok kievskiy
YieldOfPotato$Cultivar <- factor(YieldOfPotato$Cultivar)
Con <- YieldOfPotato$Cultivar == "Svitanok kievskiy" & YieldOfPotato$Year==2015
d1 <- YieldOfPotato[Con,]
anova.yieldsk<-aov(Yield.Potato~Treatment,data=d1) #anova 
anova.yieldsk.sqrt <- aov(sqrt(Yield.Potato)~Treatment,data=d1) #square root 
shapiro.test(anova.yieldsk$residuals) #0.0553
#shapiro.test(anova.yieldsk.sqrt$residuals) #0.02795

#Residuals Yield 2015 Yuna
YieldOfPotato$Cultivar <- factor(YieldOfPotato$Cultivar)
Con1 <- YieldOfPotato$Cultivar == "Yuna" & YieldOfPotato$Year==2015
d2 <- YieldOfPotato[Con1,]
anova.yieldy<-aov(Yield.Potato~Treatment,data=d2)#anova 
anova.yieldy.sqrt <- aov(sqrt(Yield.Potato)~Treatment,data=d2) #square root
shapiro.test(anova.yieldy$residuals) #0.4242
#shapiro.test(anova.yieldy.sqrt$residuals) #0.3887

#Residuals Yield 2016 Svitanok kievskiy
YieldOfPotato$Cultivar <- factor(YieldOfPotato$Cultivar)
Con1234 <- YieldOfPotato$Cultivar == "Svitanok kievskiy" & YieldOfPotato$Year==2016
d1234 <- YieldOfPotato[Con1234,]
anova.yieldsk6<-aov(Yield.Potato~Treatment,data=d1234)#anova
anova.yieldsk6.sqrt <- aov(sqrt(Yield.Potato)~Treatment,data=d1234)#square root
shapiro.test(anova.yieldsk6$residuals) #0.1674
#shapiro.test(anova.yieldsk6.sqrt$residuals) #0.1481

#Residuals Yield 2016 Yuna
YieldOfPotato$Cultivar <- factor(YieldOfPotato$Cultivar)
Con12345 <- YieldOfPotato$Cultivar == "Yuna" & YieldOfPotato$Year==2016
d2345 <- YieldOfPotato[Con12345,]
anova.yieldy6<-aov(Yield.Potato~Treatment,data=d2345) #anova
anova.yieldy6.sqrt <- aov(sqrt(Yield.Potato)~Treatment,data=d2345) #square root
shapiro.test(anova.yieldy6$residuals) #0.6108
#shapiro.test(anova.yieldy6.sqrt$residuals) #0.6299

#Graph 
par(mfrow = c(2, 2), oma = c(0, 0, 1.1, 0))
p1 <- plot(anova.yieldsk, main = "ANOVA 2015 SK", which = 2, ask = F,sub.caption = "") 
p2 <- plot(anova.yieldy, main = "ANOVA 2015 Y", which = 2, ask = F,sub.caption = "") 
p3 <- plot(anova.yieldsk6, main = "ANOVA 2016 SK", which = 2, ask = F,sub.caption = "") 
p4 <- plot(anova.yieldy6, main = "ANOVA 2016 y", which = 2, ask = F,sub.caption = "")
```

<font size="1"> Figure No. 18 Normality of Residuals Yield of Potato</font>

**3) Antifungal activity of B. velezensis against *R solani* in vitro.** 
```{r include = FALSE}
##### 1. ANTAGONISTIC ACTIVITY
#### Antagonistic activity was evaluated by measuring the diameter of the fungal 
#### colonies after 3, 5, and 7 days. The fungal inhibition (%) was determined 
#### based on the data obtained. 
#### All results were expressed and analyzed as the mean.
#### 1.1 Mean and Standard error for Fungal diameter for each treatment.
## Creating a new plot with means and standard errors
#install.packages("dplyr")
library("dplyr")
plot.means <- AntagonisticActivity %>%
  group_by(Days,Trat) %>%
  summarize(n = n(),
            mean = mean(Colony.Diameter),
            sd = sd(Colony.Diameter),
            se = sd / sqrt(n))
```


```{r}
## Graph the influence of treatments on the fungal diameter growth after 3, 5 
## and 7 days.
#install.packages("ggplot2")
library(ggplot2)
ggplot(plot.means, aes(x = Days,
                          y = mean,
                          group=Trat,
                          color=Trat)) +
  geom_point(size = 3) +
  geom_line(size = 1) +
  geom_errorbar(aes(ymin=mean-se,
                    ymax=mean+se),
                width = .1)+
  labs(title = "Graphic 1. The influence of B. velezensis strains on the growth of R. solani in vitro",
       subtitle = "(Colony diameter)",
       x = "Days",
       y = "Colony diameter",
       color = "Trat") +
  theme_light()
```

## Conclusions 


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


